---
title: ""
author: "Rodrigo Arcoverde Cerveira"
date: '`r format(Sys.Date(), "%Y-%m-%d")`'
output: html_document
abstract: Short project description.
knit: (function(inputFile, encoding) {
          rmarkdown::render(inputFile,
                            encoding = encoding, 
                            output_file = paste0(
                              xfun::sans_ext(inputFile), '_', Sys.Date(), '.html'),
                                output_dir = "~/Box Sync/public_repertoire/results")})
---

```{r setup, include=FALSE}
setwd("~/Box Sync/public_repertoire/results")
knitr::opts_knit$set(echo = TRUE,
                     root.dir = getwd(),
                     fig.width = 7, fig.height = 5,
                     warning = FALSE, 
                     message = FALSE)

result.dir <- paste0("results/",Sys.Date(),"/")

## creates result.dir with date in if not existent
ifelse(isFALSE(dir.exists(paste0("../",result.dir))), dir.create(paste0("../",result.dir),recursive = TRUE),"Result directory for today exists already!")
options(stringsAsFactors = FALSE) 
```

## Needed libraries
```{r libraries, message=FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(tidyr)
library(cowplot)
library(stringr)
```

## Load data
```{r message=FALSE, warning=FALSE, echo=FALSE}
setwd("~/Box Sync/public_repertoire/results/2020-09-13/")
ls <- list.files("public-tables", full.names = T)
ls <- grep("2020-09-13" ,ls,value = T)
ls <- grep("summary" ,ls,value = T)
ls
ls <- lapply(ls,read.csv)
names(ls) <- c("B1","B2","IgM")
df <- rbindlist(ls,idcol=T, fill = T)

#combining animals
rem_dup.one <- function(x){
  paste(unique(toupper(trimws(unlist(strsplit(x,split="(?!')[ [:punct:]]",fixed=F,perl=T))))),collapse = ", ")
}
rem_dup.vector <- Vectorize(rem_dup.one,USE.NAMES = F)

df_f <- df %>% 
  mutate(combined_animals = paste0(animals_rep_seq, sep = ", ", animals_sc)) 

df_f$combined_animals <- rem_dup.vector(df_f$combined_animals)
df_f$combined_animals <- gsub(", , ", ", ",df_f$combined_animals)
df_f$.id <- gsub("IgM", "PV", df_f$.id)
df_f$sequence <- sub("+", "",df_f$sequence)
df_comb <- df_f %>% mutate(number_animals_combined = str_count(combined_animals, "E"))
groups <- data.frame(NP = c("E11", "E16", "E17", "E23", "E24"),
                     SOL = c("E12", "E14", "E18", "E21", NA))
ff = function(x, patterns, replacements = patterns, fill = NA, ...)
{
  stopifnot(length(patterns) == length(replacements))
  
  ans = rep_len(as.character(fill), length(x))    
  empty = seq_along(x)
  
  for(i in seq_along(patterns)) {
    greps = grepl(patterns[[i]], x[empty], ...)
    ans[empty[greps]] = replacements[[i]]  
    empty = empty[!greps]
  }
  
  return(ans)
}
df_comb$NP <- ff(df_comb$combined_animals, 
                     patterns = c("E11", "E16", "E17", "E23", "E24"), 
                     replacements = c("NP","NP","NP","NP","NP"))
df_comb$SOL <- ff(df_comb$combined_animals, 
                  patterns = c("E12", "E14", "E18", "E21"), 
                  replacements = c("SOL","SOL","SOL","SOL"))

df_comb <- df_comb %>%
  mutate(Groups = ifelse(!is.na(df_comb$NP) & !is.na(df_comb$SOL),
                         "Both",
                         ifelse(!is.na(df_comb$NP) & is.na(df_comb$SOL),
                                "NP",
                                ifelse(is.na(df_comb$NP) & !is.na(df_comb$SOL),
                                       "SOL", ""))))
filtered <- df_comb %>% filter(Groups == "NP" & number_animals_combined >= 2)
filtered_sol <- df_comb %>% filter(Groups == "SOL" & number_animals_combined >= 2)
lc_all <- openxlsx::read.xlsx("~/Box Sync/RSV NGS/v4_Analysis/LCs/200828_LC_public.xlsx", sheet = 2)


```


### Plot of all public clones (left) and the ones we have light chains (right)

Public clones found in >= 2 animals across different timepoints.

```{r plot_1, warning = FALSE, fig.align= 'center',echo=FALSE, fig.width=10}
set.seed(10)
df_comb$.id <- factor(df_comb$.id ,levels = c("PV", "B1", "B2"))

#plotting with lines connecting the points
df_comb_jit <- transform(df_comb, dmeasure =  jitter(as.numeric(as.factor(.id)), 2),
jvalue = jitter(number_animals_combined, amount = .2) )

#plotting for the ones we have LCs
df_comb_jit$sequence <- gsub("_HC_PostF|_HC_PreF+|_HC_PreF|_HC_DP|_HC_DP1|_HC_DP2|+",
      "",df_comb_jit$sequence)
ptn <- paste(lc_all$well,collapse="\\b|\\b")
matches <- unique(grep(paste0("\\b", ptn,"\\b"), 
                        df_comb_jit$sequence, value=TRUE))
df_comb_jit_with_LC <- df_comb_jit[df_comb_jit$sequence %in% matches]

```


```{r data_prep, echo = FALSE, warning = FALSE}
#function to create clonoquery file first
.clono_full <- function(z){
ls <- list.files(z, full.names = T)
ls <- grep("full.txt", ls, value = T)
  if(length(ls) != 9) {
    stop("Folder not containing 9 files, please check the number of files or change the function to allow a different number of files.")
  }
  if(any(!grepl("full",ls))){
    stop("Files in the folder are not identified as a full file in its name.")
  }
  else{
    x <- lapply(ls, fread, fill = T)
    names(x) <- c("E11", "E12", "E14", "E16", "E17", "E18", "E21", "E23", "E24")
    x <- lapply(x, function (x) x[!apply(is.na(x) | x == "", 1, all),])       
    x_all <- do.call(rbind, x)
#getting the clonoquery named properly, all clones queries, aka full table
    x_m_full <- x_all %>% 
      mutate(grp = cumsum(grepl("Query",name))) %>%
      group_by(grp) %>% 
      mutate(clonoquery = ifelse(row_number() > 1, first(name), name)) %>%
      ungroup() %>% 
      select(-c(grp)) %>%
      relocate(clonoquery) %>%
      mutate(clone = substr(clonoquery, 10,14))
    }
return(x_m_full)
}

###### COMBINING CLONOQUERY FULL v4_NGS TR1 #######
clonoquery_full_B2 <- .clono_full(
  "~/Box Sync/RSV NGS/v4_Analysis/v4_public-repertoire/analysis/v4/results/2020-09-13/clonoquery/B2")

###### COMBINING CLONOQUERY FULL v4_B1-D14 #######
clonoquery_full_B1 <- .clono_full(
  "~/Box Sync/RSV NGS/v4_Analysis/v4_public-repertoire/analysis/v4/results/2020-09-13/clonoquery/B1")

###### COMBINING CLONOQUERY FULL igM #######
clonoquery_full_igm <- .clono_full(
  "~/Box Sync/RSV NGS/v4_Analysis/v4_public-repertoire/analysis/v4/results/2020-09-13/clonoquery/IgM/")

##### Combining all the clonoquery full files into a single file #######
## output as CSV file with a new column for time_point

combined_full <- list(clonoquery_full_igm, clonoquery_full_B1, clonoquery_full_B2)
names(combined_full) <- c("PV", "B1", "B2") 
combined_full  <- rbindlist(combined_full, idcol = T, fill = T)  
combined_full  <- combined_full %>% rename(time_point = .id) %>% filter(count > 0)

filtered_np <- merge(filtered, combined_full, by = "clone")
filtered_sol <- merge(filtered_sol, combined_full, by = "clone")
filtered_np <- filtered_np %>%
  mutate(E11_and_E16 = ifelse(grepl("^E11, E16$",combined_animals),"YES", "NO"))

```


###  Nanoparticle group (only) by specificity, clone size, and animals

The graph on the left includes all of the clones that are within the nanoparticle, while the graph on the right shows only the clones light chains are available. 


```{r plot_6, echo = FALSE, warning = FALSE, fig.align= 'center', fig.width=20,fig.height= 10} 
filtered_np_spec <- filtered_np %>%  mutate(
  spec = ifelse(grepl("(?=.*DP)(?=.*PreF)", sequence, perl = T), "Both",
                ifelse(grepl("PreF",sequence, perl = T), "PreF", "DP"))) %>%
  group_by(clone) %>% 
  summarise_each(funs(if(is.numeric(.)) mean(., na.rm = TRUE) else dplyr::first(.))) %>% 
  mutate(V_gene = gsub("-.*","", V_gene),
         clone_percentage = (size_rep_seq/sum(size_rep_seq))*100) %>% 
  as.data.frame() 

filtered_sol_spec <- filtered_sol %>%  mutate(
  spec = ifelse(grepl("(?=.*DP)(?=.*PreF)", sequence, perl = T), "Both",
                ifelse(grepl("PreF",sequence, perl = T), "PreF", "DP"))) %>%
  group_by(clone) %>% 
  summarise_each(funs(if(is.numeric(.)) mean(., na.rm = TRUE) else dplyr::first(.))) %>% 
  mutate(V_gene = gsub("-.*","", V_gene),
         clone_percentage = (size_rep_seq/sum(size_rep_seq))*100) %>% 
  as.data.frame() 
p10 <- filtered_sol_spec %>% 
  ggplot(aes(x = V_gene, y = clone_percentage, fill = spec)) +
  geom_col(size = .1, width = .5, color = "Black") +
  scale_fill_viridis_d() +
  theme_cowplot() +
  labs(x = "", y = "Clone size (%)", fill = "Specificity", title = "Public soluble repertoire")
p10
p11 <- filtered_np_spec %>% 
  ggplot(aes(x = V_gene, y = clone_percentage, fill = spec)) +
  geom_col(size = .1, width = .5, color = "Black") +
  scale_fill_viridis_d() +
  theme_cowplot() +
  labs(x = "", y = "Clone size (%)", fill = "Specificity", title = "Public nanoparticle repertoire")

combined_v_gene <- merge(df_comb, combined_full, by = "clone")
combined_full_v_gene <- combined_v_gene %>%  mutate(
  spec = ifelse(grepl("(?=.*DP)(?=.*PreF)", sequence, perl = T), "Both",
                ifelse(grepl("PreF",sequence, perl = T), "PreF", "DP"))) %>%
  group_by(clone) %>% 
  summarise_each(funs(if(is.numeric(.)) mean(., na.rm = TRUE) else dplyr::first(.))) %>% 
  mutate(V_gene = gsub("-.*","", V_gene),
         clone_percentage = (size_rep_seq/sum(size_rep_seq))*100) %>% 
  as.data.frame() 



p12 <- combined_full_v_gene  %>% 
  ggplot(aes(x = V_gene, y = clone_percentage, fill = spec)) +
  geom_col(size = .1, width = .5, color = "Black") +
  scale_fill_viridis_d() +
  theme_cowplot() +
  labs(x = "", y = "Clone size (%)", fill = "Specificity", title = "Total public repertoire")


table_query <- read.csv("~/Box Sync/RSV NGS/v4_Analysis/v4_public-repertoire/merged_sc_clonotypes/2020-08-10_public_clonotypes_query_ref-table.csv", header = T) %>% rename(clone = Public.clonotype)




comb_query <- merge(combined_full, table_query, by = "clone") %>% 
  mutate(spec = ifelse(grepl("(?=.*DP)(?=.*PreF)", name.y, perl = T), "Both",
                ifelse(grepl("PreF",name.y, perl = T), "PreF", "DP")))

##### add the clones that have both spec manually #####
x <- combined_full_v_gene$clone[combined_full_v_gene$spec == "Both"]
comb_query <- comb_query %>% mutate(spec = ifelse(clone %in% x,
                                    "Both",
                                    spec))


p13 <- comb_query  %>% group_by(clone) %>% 
  summarise_each(funs(if(is.numeric(.)) mean(., na.rm = TRUE) else dplyr::first(.))) %>% 
  mutate(V_gene = gsub("-.*","", V_gene.x),
         clone_percentage = (count.x/sum(count.x))*100) %>% 
  as.data.frame() 


p14 <- p13 %>%  ggplot(aes(x = V_gene, y = clone_percentage,  fill = spec)) +
  geom_col(size = .1, width = .5, color = "Black") + scale_fill_viridis_d() + theme_cowplot() +
  labs(x = "", y = "Clone size (%)", title = "Total repertoire", fill = "Specificity")


gridExtra::grid.arrange(p14, p12, p11, p10, nrow = 1)
gridExtra::grid.arrange(p14, p12, nrow = 1)

```

## SessionInfo
```{r}
sessionInfo()
```

